---
title: 'Activity: Popular Names'
author: "name"
date: "June 3, 2024"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---


# Grading

Assignment is worth a total of 10 points

* [1 point] Turn in BOTH RNotebook (.nb.html) and the RMarkdown (.Rmd) file to Canvas
* [1 point] Changes pushed to GitHub
* [1 point] Step 1: show how you used summary functions to explore the raw data set
* [1 point] Step 3 and 4: attempt to answer all questions about the data
* [1 point] Step 5: explanation of data wrangling steps in English (not code)
* [3 points] Step 7: Data wrangling
    * correct use of `filter()`
    * correct use of `group_by()`
    * correct use of `summarise()`
* [2 points] `ggplot()` graph of first names 



# Set Up

```{r, message = F, warning = F}
# load necessary packages
library(tidyverse)
library(mosaic)
library(esquisse)

devtools::install_github("mdbeckman/dcData")
library(dcData)

# Load the BabyNames data into RStudio environment (from dcData package)
data("BabyNames", package = "dcData")
```

# Activity

**Goal**: By the end of the activity, each team will create a graph (like Figure A.1 on p 159) showing ups and downs in the popularity of the names at your table according to the U.S. social security registry.  

For steps 5-7, pick any FIVE names you wish. I choose Arnaldo, Ahmid, Nicho, Jennaveve, and Christy. You may choose these names or any others that you would like. 


### Step 1

*Examine the data you have at hand - for this project the data are the ones available in the `BabyNames` data set. Use at least 2 different functions to examine your data. Then answer the questions: What are the variables are available? What is a case?*

##### What are the variables are available?


##### What is a case?
 
## Look at the data a few different ways (see Ch 9 for more).
```{r}


```

### Step 2

[This should be done by hand on paper, nothing to show in the HTML doc]


### Step 3 

##### What variable(s) from the raw data table do not appear at all in the graph?

##### What variable(s) in the graph are similar to corresponding variables in the raw data table, but might have been transformed in some way?



### Step 4

##### Have cases been filtered out?



##### Have cases been grouped and summarized within groups in any way?


##### Have any new variables been introduced?  If so, what the relationship between the new variables and existing variables?



### Step 5
Using the BabyNames data set, filter the data to display just the 5 names that you picked.  Then group the cases by year and name, and count how many total births per year were recorded for each of the names (regardless of sex).  

1. Select the five names

2. Filter the BabyNames dataset for only those 5 names selected in Step-1 and store the output data in separate variable as a dataframe

3. Use "group_by" function to group by the output data from Step-2 at "name" and "year" level

4. Use "summarise" function to get the total number of births  recorded for each of the names per year on the grouped by data from Step-3


### Step 6

[This should be done by hand on paper, nothing to show in the HTML doc]

### Step 7
#### Create my glyph ready data set here
```{r}



```


### Graph It!

Here's the sample code provided by the book:

```
GlyphReadyForm %>%
  ggplot(aes(x = year, y = total, group = name)) + 
  geom_line(size = 1, alpha = 0.5, aes(color = name)) + 
  ylab("Popularity") + 
  xlab("Year") +
  ggtitle("Popularity of Names Over Time")
```

If you prepared your `GlyphReadyForm` data set correctly, you should just be able to copy and paste this code. Alternatively, you can use `mplot` or `esquisse` to create your graph. If you choose to use these interactive functions remember, do not include the `mplot` or `esquisse` functions in your RMarkdown document. Instead, use these functions in the console and copy the generated code into your RMarkkdown document. In addition, we want lines, not dots.




